---
datatitle: "Guinea Pig Tooth Growth"
author: "Ralph Fehrer"
date: "Sunday, November 23, 2014"
output: pdf_document
---
###1.Load the ToothGrowth data and perform some basic exploratory data analyses 

The R ToothGrowth data set contains data from an experiment studying the effect of vitamin C on the tooth growth of guinea pigs. 
```{r,echo=TRUE}
library(datasets)
data(ToothGrotwh)
```
The vitamin C was supplied through two different supplements, orange juice and asorbic acid
```{r,echo=TRUE}
str(ToothGrowth$supp)
```
in three doses each
```{r,echo=TRUE}
str(as.factor(ToothGrowth$dose))
```
###2. Provide a basic summary of the data.
```{r,echo=TRUE}
summary(ToothGrowth)
library(ggplot2)
ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose)))+geom_boxplot()+facet_grid(.~supp)+ggtitle("Effect of vitamin C on tooth growth of guinea pigs")
```
###3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose
```{r}
xBar<-mean(ToothGrowth$len[1:30])
yBar<-mean(ToothGrowth$len[31:60])
xVar<-(sd(ToothGrowth$len[1:30]))^2
yVar<-(sd(ToothGrowth$len[31:60]))^2
q<-(((xVar+yVar)/30)^2)/((((xVar/30)^2)+((yVar/30)^2))/29)
t<-qt(0.975, q)
yBar -xBar + c(-1,1)*t*sqrt(xVar/30 + yVar/30)

```
```{r}
t.test(len~supp, data=ToothGrowth, paired=FALSE)
```
###4.State your conclusions and the assumptions needed for your conclusions
the boxplot shows, that vitamin C has an effect on teeth growth in guinea pigs. The effect increases as the dose increases, regardless of the supplement. 

