---
title: "StatisticalInferenceProject"
author: "Ralph Fehrer"
date: "Wednesday, November 19, 2014"
output: pdf_document
---

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
For the experiments in this paper, lambda is set to 

```{r}
lambda = 0.2
```

The mean and the standard deviation of an exponential distribution are 

```{r}
exponentialMean = exponentialSd = 1/lambda
```





for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. 

```{r}
numberOfDistributionSamples = 40
```

Note that you will need to do a thousand or so simulated averages of 40 exponentials.
```{r}
numberOfParameterSamples = 1000
```

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s.  You should


```{r}
numberOfDistributionSamples
```


3. Show that the distribution is approximately normal.

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.

```{r, echo=FALSE}
mns = NULL
for (i in 1 : numberOfParameterSamples) mns = c(mns, mean(rexp(numberOfDistributionSamples,lambda)))
hist(mns)
realMean=1/lambda
sampleMean=mean(mns)
differenceBetweenSampleAndTheoreticalMean=abs(exponentialMean-sampleMean)
```

2. Show how variable it is and compare it to the theoretical variance of the distribution.

```{r, echo=FALSE}
sds = NULL
for (i in 1 : numberOfParameterSamples) sds = c(sds, sd(rexp(numberOfDistributionSamples,lambda)))
hist(sds)
sampleMean=mean(sds)
differenceBetweenSampleAndTheoreticalSd=abs(exponentialSd-sampleMean)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
